---
title: "Adv Econ II Assignment 5"
author: "A. Schmidt"
date: "2/9/2020"
output: html_document
---

# Loading packages

```{r}
list.of.packages <- c("ggplot2", "latex2exp")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
library(ggplot2, quietly = TRUE)
library(latex2exp, quietly = T)
```

# Generating data

We are considering the following DGP:

$$y_{it} = \beta x_{it} + \alpha_i + \varepsilon_{it}$$
where $\alpha_{i} \sim \mathcal{N}(0, \sigma^2_a)$, $\varepsilon_{i} \sim \mathcal{N}(0, \sigma^2_\varepsilon)$. The regressor $x_{it}$ follows an AR(1) process given by

$$x_{it} = \rho x_{i, t-1} + \pi \alpha_i + \theta \varepsilon_{i,t-1} + \xi_{it},$$
where $\xi_{it} \sim \mathcal{N}(0, \sigma^2_{\xi})$. The process starts at

$$x_{i0} = \frac{\pi \alpha_i}{1-\rho} + \frac{\theta \nu + \zeta}{\sqrt{1-\rho^2}},$$
where both $\nu$ and $\zeta$ are i.i.d. following a standard normal distribution.

## Magic Numbers

The following variables must be entered by the user:

* `iNobs` - Integer, the number of observations. Default value is 1000.
* `dSigmaalpha` - Double, the variance of the random effect $\alpha_{i}$. Default value is 1.
* `dSigmaepsilon` - Double, the variance of the error term from the regression equation, $\varepsilon_{it}$. Default value is 1.
* `dSigmaxi` - Double, the variance of the random error term $\xi_{it}$ in the autoregressive process for $x_{it}$. Default value is 1.]
* `dBeta` - Double, the parameter multiplying $x_{it}$ in the regression equation. Defailt value is 1.
* `dRho` - Double, persistence parameter of the AR(1) process for $x_{it}$. Defailt value is 0.5.
* `dPì` - Double, the coefficient of $\alpha_i$ in the autoregressive process for $x_{it}$. Notice that when $\pi \neq 0$, we no longer have a random effect in the regression equation. Default value is 1.
* `dTheta` - Double, the coefficient multiplying $\varepsilon_{i, t-1}$ in the $x_{it}$ equation. When not zero, it introduces autocorrelation of the errors in the regression equation. Default value is 0. When zero, we have strong exogeneity of $x_{it}$.


```{r}
iNobs          <- 10
dSigmaalpha    <- 1
dSigmaepsilon  <- 1
dSigmaxi       <- 1
dBeta          <- 1
dRho           <- 0.5
dPi            <- 1
dTheta         <- 0

# Set a seed
set.seed(6969)
```

## Generating data functions

### Errors

Errors are generating drawing from a normal distribution, where the number of observations as well as the mean and variance can be changed when calling the function.

```{r}
fGenError <- function(iNobs, dMu = 0, dSigma2 = 1)
{
  # Returns a vector with length iNobs with i.i.d observations
  # from a normal density with mean dMu and variance dSigma2
  vdError <- rnorm(iNobs, dMu, dSigma2)
  return(vdError)
}
```


### X

The X variables follows an AR(1) process with a predetermined X0.

```{r}
fGenX <- function(iNobs, dPi = 1, vdAlpha, dRho = 0.5, dTheta = 0)
{
  # Generates X using a predetermined first element and an autoregressive process for the others.
  dNu    <- fGenError(1)
  dZeta  <- fGenError(1)
  dXi0   <- ((dPi*vdAlpha[1])/(1-dRho))+ ((dTheta*dNu+dZeta)/(sqrt(1-dRho^2)))
  vdX     <- rep(0, iNobs)
  vdX[1]  <- dXi0

  for(i in 2:iNobs){
    vdX[i] <- dRho * vdX[i-1] + dPi * vdAlpha[i] + dTheta * vdEpsilon[i-1] + vdXi[i]
  }
  
  return(vdX)
}
```

### Y

Generating $y$ requires to have $\beta$, $x_{it}$, $\alpha_i$ and $\varepsilon_{it}$ to assemble the regression.

```{r}
fGenY <- function(dBeta = 1, vdX, vdAlpha = 0, vdEpsilon){
  
  vdY = dBeta * vdX + vdAlpha + vdEpsilon
  
  return(vdY)
}
```

### Instruments

Three instruments' matrices are used:

\begin{equation}
Z_{1i} = 
\begin{bmatrix}
x_{i1} \\ 
x_{i2} \\
\vdots \\
x_{i, T-1}
\end{bmatrix} \qquad
Z_{2i} = 
\begin{bmatrix}
x_{i1} & 0 & \cdots & 0 \\ 
0 & x_{i2} & & \vdots \\
\vdots & & \ddots & 0 \\
0 & \cdots & 0 & x_{i, T-1}
\end{bmatrix} \qquad
Z_{3i} = 
\begin{bmatrix}
x_{i1} & 0 & 0 & \cdots & 0 & \cdots & 0 \\ 
0 & x_{i1} & x_{i2} & \cdots & 0 & \cdots & 0 \\
.& . & . & \cdots & . & \cdots & . \\
0 & 0 & 0 &  \cdots & x_{i1} & \cdots &  x_{i, T-1}
\end{bmatrix}
\end{equation}

```{r}

  fGenInstruments <- function(vdX, inst){
    dNum <- length(vdX[1:(length(vdX)-1)])
    
    if (inst == 1){
      
      mdZ <- vdX[1:dNum]  
    
      } else if (inst == 2){
      
        mdZ <- diag(vdX[1:dNum])  
    
        } else {
      
          mdZ <- matrix(0, nrow = dNum, ncol = (0.5*dNum*(dNum-1)+dNum))
      
          for (i in 1:nrow(mdZ)){
        
            if (i == 1){
          
              mdZ[1,] <- c(vdX[1], rep(0, (0.5*dNum*(dNum-1))+dNum-1))
        
              } else {
          
                mdZ[i,] <- c(rep(0, (0.5*i*(i-1))), vdX[1:i], rep(0, ((0.5*dNum*(dNum-1)+dNum))-length(c(rep(0, (0.5*i*(i-1))), vdX[1:i]))))
                }
            }  
        }
    return(mdZ)
    }
```

### Generating the data

```{r}
# Errors 
  vdAlpha    <- fGenError(iNobs, 0, dSigmaalpha)
  vdEpsilon  <- fGenError(iNobs, 0, dSigmaepsilon)
  vdXi       <- fGenError(iNobs, 0, dSigmaxi)

# X
  vdX <- fGenX(iNobs, dPi, vdAlpha, dRho, dTheta)

# Y
  vdY <- fGenY(dBeta, vdX, vdAlpha, vdEpsilon)

# Instruments
  mdZ1 <- fGenInstruments(vdX, 1)
  mdZ2 <- fGenInstruments(vdX, 2)
  mdZ3 <- fGenInstruments(vdX, 3)

# Lagged Variables
  vdXlag  <- c(NA, vdX[seq_along(vdX) -1])
  vdYlag  <- c(NA, vdY[seq_along(vdY) -1])
  vdXdiff <- vdX[2:length(vdX)] - vdXlag[2:length(vdXlag)]
  vdYdiff <- vdY[2:length(vdY)] - vdYlag[2:length(vdYlag)]
```

### Estimators

#### OLS

The OLS estimator is given by $\hat{\beta} = (X'X)^{-1}X'Y$ with standard error given by $\hat{\sigma}^2_{\varepsilon}(X'X)^{-1}$.

##### Regular OLS:

```{r}
  fEstOLS <- function(vdX, vdY) {
    vdBetaOLS <- solve((t(vdX)%*%vdX))%*%(t(vdX)%*%vdY)
  }
```

##### Pooled OLS:

```{r}
  fEstPOLS <- function(vdX, vdY) {
    
  }
```

##### First Difference OLS:

```{r}
  fEstFDOLS <- function(vdX, vdY) {
    
  }
```

##### Whithin OLS:

```{r}
  fEstFEOLS <- function(vdX, vdY) {
    
  }
```

#### Instrumental variables estimator

```{r}
  fEstIV <- function(vdX, vdY, mdZ) {
    
  }
```

#### GMM

#### Instrumental variables estimator

```{r}
  fEstGMM <- function(vdX, vdY, mdZ) {
    
  }

```