---
title: "Adv Econ II Assignment 5 - Programming"
author: "A. Schmidt, E. Ugulava, X. Lin"
date: "02/10/2020"
output: html_document
---

To make references: bibliography: references.bib


# Loading packages

```{r, warning = FALSE, message = FALSE}
list.of.packages <- c("ggplot2", "latex2exp")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)
library(ggplot2, quietly = TRUE)
library(latex2exp, quietly = T)
```

# Generating data 

We are considering the following DGP:

$$y_{it} = \beta x_{it} + \alpha_i + \varepsilon_{it}$$
where $\alpha_{i} \sim \mathcal{N}(0, \sigma^2_a)$, $\varepsilon_{i} \sim \mathcal{N}(0, \sigma^2_\varepsilon)$. The regressor $x_{it}$ follows an AR(1) process given by

$$x_{it} = \rho x_{i, t-1} + \pi \alpha_i + \theta \varepsilon_{i,t-1} + \xi_{it},$$
where $\xi_{it} \sim \mathcal{N}(0, \sigma^2_{\xi})$. The process starts at

$$x_{i0} = \frac{\pi \alpha_i}{1-\rho} + \frac{\theta \nu + \zeta}{\sqrt{1-\rho^2}},$$
where both $\nu$ and $\zeta$ are i.i.d. following a standard normal distribution.

## Magic Numbers

The following variables must be entered by the user:

* `iNobs` - Integer, the number of observations. Default value is 100.
* `iTime` - Integer, the number of time periods in which we observe each individual. Default is 5.
* `iIter` - Integer, the number of iterations for the simulation. Default is 1000.
* `dSigmaalpha` - Double, the variance of the random effect $\alpha_{i}$. Default value is 1.
* `dSigmaepsilon` - Double, the variance of the error term from the regression equation, $\varepsilon_{it}$. Default value is 1.
* `dSigmaxi` - Double, the variance of the random error term $\xi_{it}$ in the autoregressive process for $x_{it}$. Default value is 1.]
* `dBeta` - Double, the parameter multiplying $x_{it}$ in the regression equation. Defailt value is 1.
* `dRho` - Double, persistence parameter of the AR(1) process for $x_{it}$. Defailt value is 0.5.
* `dPì` - Double, the coefficient of $\alpha_i$ in the autoregressive process for $x_{it}$. Notice that when $\pi \neq 0$, we no longer have a random effect in the regression equation. Default value is 1.
* `dTheta` - Double, the coefficient multiplying $\varepsilon_{i, t-1}$ in the $x_{it}$ equation. When not zero, it introduces autocorrelation of the errors in the regression equation. Default value is 0. When zero, we have strong exogeneity of $x_{it}$.


```{r}
iNobs          <- 10
iTime          <- 3
iIter          <- 1
dSigmaalpha    <- 1
dSigmaepsilon  <- 1
dSigmaxi       <- 1
dBeta          <- 1
dRho           <- 0.5
dPi            <- 1
dTheta         <- 0

# Set a seed
set.seed(6969)
```

## Generating data functions

### Errors

Errors are generating drawing from a normal distribution, where the number of observations as well as the mean and variance can be changed when calling the function.

```{r}
fGenError <- function(iTime, dMu = 0, dSigma2 = 1)
{
  # Returns a vector with length iNobs with i.i.d observations
  # from a normal density with mean dMu and variance dSigma2
  vdError <- rnorm(iTime, dMu, dSigma2)
  return(vdError)
}
```


### X

The X variables follows an AR(1) process with a predetermined X0.

```{r}
fGenX <- function(iTime, dPi = 1, dAlpha, dRho = 0.5, dTheta = 0, vdEpsilon, vdXi)
{
  # Generates X using a predetermined first element and an autoregressive process for the others.
  dNu    <- fGenError(1)
  dZeta  <- fGenError(1)
  dXi0   <- ((dPi*dAlpha)/(1-dRho))+ ((dTheta*dNu+dZeta)/(sqrt(1-dRho^2)))
  vdX     <- rep(0, iTime)
  vdX[1]  <- dXi0

  for(i in 2:iTime){
    #print(i)
    vdX[i] <- dRho * vdX[i-1] + dPi * dAlpha + dTheta * vdEpsilon[i-1] + vdXi[i]
  }
  
  return(vdX)
}
```

### Y

Generating $y$ requires to have $\beta$, $x_{it}$, $\alpha_i$ and $\varepsilon_{it}$ to assemble the regression.

```{r}
fGenY <- function(dBeta = 1, vdX, dAlpha = 0, vdEpsilon){
  
  vdY = dBeta * vdX + rep(dAlpha, iTime) + vdEpsilon
  
  return(vdY)
}
```

### Whole panel

Generating the panel requires that for each individual we simulate the three errors, the $X$ variable and put together with $\beta$ to create $Y$.

```{r}
fGenPanel <- function(iNobs = 100, iTime = 5, dPi = 1, dRho = 0.5, dTheta = 0, dBeta = 1, dSigmaalpha = 1, dSigmaepsilon = 1, dSigmaxi = 1) {
  for (i in 1:iNobs){
    # Errors 
    dAlpha     <- fGenError(1, 0, dSigmaalpha)
    vdEpsilon  <- fGenError(iTime, 0, dSigmaepsilon)
    vdXi       <- fGenError(iTime, 0, dSigmaxi)

    # X
    vdX <- fGenX(iTime, dPi, dAlpha, dRho, dTheta, vdEpsilon, vdXi)

    # Y
    vdY <- fGenY(dBeta, vdX, dAlpha, vdEpsilon)
    
    # Lagged Variables
    vdXlag  <- c(NA, vdX[seq_along(vdX) -1])
    vdYlag  <- c(NA, vdY[seq_along(vdY) -1])
    vdXdiff <- vdX - vdXlag
    vdYdiff <- vdY - vdYlag
      
    # Assembling the panel
  
    if (i == 1) {
      dfPanel <- data.frame(cbind(rep(i, iTime), 1:iTime, rep(dAlpha, iTime), vdEpsilon, vdXi, vdX, vdY, vdXdiff, vdYdiff))
    } else {
      dfPanel <- data.frame(rbind(dfPanel, cbind(rep(i, iTime), 1:iTime, rep(dAlpha, iTime), vdEpsilon, vdXi, vdX, vdY, vdXdiff, vdYdiff)))
    }
  }
  names(dfPanel) <- c("Individual", "Time", "Alpha", "Epsilon", "Xi", "X", "Y", "Xdiff", "Ydiff")

  return(dfPanel)
}
```

### Instruments

Three instruments' matrices are used:

\begin{equation}
Z_{1i} = 
\begin{bmatrix}
x_{i1} \\ 
x_{i2} \\
\vdots \\
x_{i, T-1}
\end{bmatrix} \qquad
Z_{2i} = 
\begin{bmatrix}
x_{i1} & 0 & \cdots & 0 \\ 
0 & x_{i2} & & \vdots \\
\vdots & & \ddots & 0 \\
0 & \cdots & 0 & x_{i, T-1}
\end{bmatrix} \qquad
Z_{3i} = 
\begin{bmatrix}
x_{i1} & 0 & 0 & \cdots & 0 & \cdots & 0 \\ 
0 & x_{i1} & x_{i2} & \cdots & 0 & \cdots & 0 \\
.& . & . & \cdots & . & \cdots & . \\
0 & 0 & 0 &  \cdots & x_{i1} & \cdots &  x_{i, T-1}
\end{bmatrix}
\end{equation}

```{r} 

  fGenInstruments <- function(vdX, inst){
    dNum <- length(vdX[1:(length(vdX)-1)])
    
    if (inst == 1){
      
      mdZ <- vdX[1:dNum]  
    
      } else if (inst == 2){
      
        mdZ <- diag(vdX[1:dNum])  
    
        } else {
      
          mdZ <- matrix(0, nrow = dNum, ncol = (0.5*dNum*(dNum-1)+dNum))
      
          for (i in 1:nrow(mdZ)){
        
            if (i == 1){
          
              mdZ[1,] <- c(vdX[1], rep(0, (0.5*dNum*(dNum-1))+dNum-1))
        
              } else {
          
                mdZ[i,] <- c(rep(0, (0.5*i*(i-1))), vdX[1:i], rep(0, ((0.5*dNum*(dNum-1)+dNum))-length(c(rep(0, (0.5*i*(i-1))), vdX[1:i]))))
                }
            }  
        }
    return(mdZ)
    }
```

#### First difference matrix

We can build a matrix $D$ that takes the first difference when multiplied by a vector. For example, if $T=3$, the matrix $D$ will be given by:

\begin{equation}
D = 
  \begin{bmatrix}
  -1 & 1 & 0 \\
  0 & -1 & 1 \\
  \end{bmatrix}.
\end{equation}

Then, the following happens:

\begin{equation}
D \cdot X_i =  
  \begin{bmatrix}
  -1 & 1 & 0 \\
  0 & -1 & 1 \\
  \end{bmatrix} \cdot   
  \begin{bmatrix}
  x_{i1} \\
  x_{i2} \\
  x_{i3}
  \end{bmatrix}
  =
  \begin{bmatrix}
  x_{i2} - x_{i1} \\
  x_{i3} - x_{i2}
  \end{bmatrix}
\end{equation}

```{r}

    fGenD <- function(iTime) {
          viD <- c(-1, 1)
          miD <- matrix(0, nrow = iTime-1, ncol = iTime)
   
          for (i in 1:(iTime-1)){
            miD[i, i:(i+1)] <- viD
          }
          
          return(miD)
    }

```


#### Weighting matrix

  The weighting matrix is given by
  
  $$W_N = \left(\sum_{i=1}^N Z_i'DD'Z_i \right)^{-1}.$$

```{r}
  fGenWeights <- function(iTime, ldZ) {
      
      mdSum <- 0
      miD <- fGenD(iTime)
  
      for (i in ldZ){
        mdSum <- mdSum + t(i) %*% miD %*% t(miD) %*% i
      }
      
      mdWN <- solve(mdSum)
      
      return(mdWN)
  }
```

### Generating the data

Panel data is different from simply making a regression because we will have repeated $i$ and $t$. In here, for each individual $i$ we create $\alpha_i$, $\varepsilon_i$ and $\xi_i$ inside a loop and put everything together into a data frame.

```{r}
  # Generates the panel
  dfPanel <- fGenPanel(iNobs, iTime, dPi, dRho, dTheta, dBeta, dSigmaalpha, dSigmaepsilon, dSigmaxi)

  # Creates lists with instruments for each individual
  ldZ1 <- list(1:iNobs)
  ldZ2 <- list(1:iNobs)
  ldZ3 <- list(1:iNobs)
  
  # This should be computed using X or Xdiff?
  for (i in 1:iNobs){
    ldZ1[[i]] <- fGenInstruments(dfPanel$X[dfPanel$Individual == i], 1)
    ldZ2[[i]] <- fGenInstruments(dfPanel$X[dfPanel$Individual == i], 2)
    ldZ3[[i]] <- fGenInstruments(dfPanel$X[dfPanel$Individual == i], 3)
  }

```

### Estimators

#### OLS

The OLS estimator is given by $\hat{\beta} = (X'X)^{-1}X'Y$ with standard error given by $\hat{\sigma}^2_{\varepsilon}(X'X)^{-1}$.

##### Pooled OLS

We use the formula from [@cameron2005] for the Pooled OLS (adapted for the fact that we have only one explanatory variable):

\begin{equation}
  \hat{\beta}_{\text{POLS}} = \left[\sum_{i=1}^N \sum_{t=1}^T x_{it}^2 \right]^{-1} \sum_{i=1}^N \sum_{t=1}^T x_{it}y_{it}
\end{equation}

```{r}
  fEstPOLS <- function(dfPanel, iNobs, iTime) {
    
    dX2sum <- 0
    dXYsum <- 0
    
    for (i in iNobs){
      for (j in iTime){
        dX2sum <- dX2sum + (dfPanel$X[dfPanel$Individual == i, dfPanel$Time == j])^2
        dXYsum <- dXYsum + (dfPanel$X[dfPanel$Individual == i, dfPanel$Time == j])*(dfPanel$Y[dfPanel$Individual == i, dfPanel$Time == j])
      }
    }
    
    dBetaPOLS <- (1/dX2sum)*dXYsum
    
    return(dBetaPOLS)
  }
```

##### First Difference OLS:

```{r, EVAL = FALSE}
  fEstFDOLS <- function(vdX, vdY) {
    
  }
```

##### Whithin OLS:

We use the formula from [@cameron2005] for the Whithin or Fixed Effects OLS (page 726):

\begin{equation}
  \hat{\beta}_{\text{FEOLS}} = \left[\sum_{i=1}^N \sum_{t=1}^T (x_{it}-\bar{x}_i)^2 \right]^{-1} \sum_{i=1}^N \sum_{t=1}^T (x_{it}-\bar{x}_i)(y_{it}-\bar{y}_i)
\end{equation}

TBC

```{r, eval = FALSE}
  fEstFEOLS <- function(dfPanel, iNobs, iTime) {
    mdQN <- 
  }
```

#### Instrumental variable estimator / GMM (?)

For the first instrument we have exact identification so it is possible to compute $\hat{\beta}_{\text{IV}}$ by using $(Z'X)^{-1}Z'Y$. For the other two instruments, we need to minimize the criteria function

$$Q_N(\beta)= g_N'(\beta)W_N g_N(\beta).$$

```{r}
  fEstIV <- function(dfPanel, iTime, ldZ, type) {
    if (type == 1){ # For the first instrument (column instrument)
      vdBetaIV <- solve(t(sapply(sapply(ldZ, rbind), rbind)) %*% dfPanel$Xdiff[!is.na(dfPanel$Xdiff)]) %*% t(sapply(sapply(ldZ, rbind), rbind)) %*% dfPanel$Ydiff[!is.na(dfPanel$Ydiff)]
      
    } else if (type %in% c(2,3)) { # For the second and third instrument
      
      mdWN <- fGenWeights(iTime, ldZ)
    
      # Stack the instruments
      mdZ <- NA
      for (i in ldZ){
        mdZ <- rbind(mdZ, i)
      }
      mdZ <- mdZ[-1,]
    
      # Stack Xdiff
      vdX <- dfPanel$Xdiff[!is.na(dfPanel$Xdiff)]
      
      # Stack Ydiff
      vdY <- dfPanel$Ydiff[!is.na(dfPanel$Ydiff)]
      
      # Criteria function
      fGenQN <- function(dBetahat) (t(t(mdZ) %*% (vdY - vdX * dBetahat)) %*% mdWN %*% (t(mdZ) %*% (vdY - vdX * dBetahat)))
      
      # Optimization
      vdBetaIV <- optimize(fGenQN, interval = c(-5,5), tol = 0.00001)$minimum

    } else {
      vdBetaIV <- 0
      print("Invalid instrument")
    }
    
      return(vdBetaIV)  
  }
```

# Estimating stuff

```{r}
  # This computes the BetaIV
  dBetaIV1 <- fEstIV(dfPanel, iTime, ldZ1, type=1)
  dBetaIV2 <- fEstIV(dfPanel, iTime, ldZ2, type=2)
  dBetaIV3 <- fEstIV(dfPanel, iTime, ldZ3, type=3)

```

# References
