---
title: "Adv. Microeconometrics Computer Assignment"
author: "A. Schmidt"
date: "11/19/2020"
output:
  pdf_document: default
bibliography: references.bib
---



```{r, message = FALSE, warning = FALSE, echo = FALSE}
# Verify if a package is already installed, if not, download and install before loading. 
chooseCRANmirror(graphics = FALSE, ind = 10)
if (!require("pacman")) install.packages("pacman")
pacman::p_load(ggplot2, kableExtra, stargazer, xts, knitr, tibble, broom, forecast, lmtest, sweep, reshape, gridExtra, ggpubr, tseries)
```

```{r, message = FALSE, warning = FALSE, echo = FALSE}
# Prevents code from getting out of the page
## Works with almost everything except urls and strings.
opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE)
```

# 1 - Size distortions

Simulate data from the following model:

\begin{align*}
  Y &= X \beta + \varepsilon \\
  X &= Z\Pi + V
\end{align*}
where:
* $Y$ and $X$ are $n \times 1$ vectors which contain the endogenous variables;
* $Z$ is a $n \times k$ matrix of instruments;
* $\varepsilon$ and $V$ are $n \times 1$ vectors that contain disturbances.

The different rows of $\left(\varepsilon\ \vdots\ V\right)$, are independently normally distributed, i.e., 
\begin{align*}
  \begin{pmatrix}
    \varepsilon_i \\
    V_i
  \end{pmatrix} \sim \mathcal{N}(0, \Sigma), \qquad 
  \Sigma =
  \begin{pmatrix}
  1 & \vdots & \rho \\
  \rho & & 1
  \end{pmatrix}
\end{align*}


