---
title: "Adv. Microeconometrics Computer Assignment"
author: "A. Schmidt"
date: "11/19/2020"
output:
  html_document:
    df_print: paged
bibliography: references.bib
---

  pdf_document: default


```{r, message = FALSE, warning = FALSE, echo = FALSE}
# Verify if a package is already installed, if not, download and install before loading. 
chooseCRANmirror(graphics = FALSE, ind = 10)
if (!require("pacman")) install.packages("pacman")
pacman::p_load(ggplot2, MASS, knitr, ivreg, gridExtra, RColorBrewer, latex2exp, ggExtra, grid)
```

```{r, message = FALSE, warning = FALSE, echo = FALSE}
# Prevents code from getting out of the page
## Works with almost everything except urls and strings.
opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE)
```

# 1 - Size distortions

Simulate data from the following model:

\begin{align*}
  Y &= X \beta + \varepsilon \\
  X &= Z\Pi + V
\end{align*}
where:
* $Y$ and $X$ are $n \times 1$ vectors which contain the endogenous variables;
* $Z$ is a $n \times k$ matrix of instruments;
* $\varepsilon$ and $V$ are $n \times 1$ vectors that contain disturbances.
* The different rows of $\left(\varepsilon\ \vdots\ V\right)$, are independently normally distributed, i.e., 

\begin{align*}
  \begin{pmatrix}
    \varepsilon_i \\
    V_i
  \end{pmatrix} \sim \mathcal{N}(0, \Sigma), \qquad 
  \Sigma =
  \begin{pmatrix}
  1 & \vdots & \rho \\
  \rho & & 1
  \end{pmatrix}
\end{align*}
* $n=100$, $k=10$, $\Pi = a \times e_{10}$ with $e_{10} \in \mathbb{R}^{10}$ whose first element is $1$ and the remaining are equal to zero.
* All elements from $Z$ are independently distributed and follows a standard normal distribution.
* $a \in \{0.3,.0.25, 0.2, 0.15, 0.1, 0.05, 0 \}$
* $\rho \in \{0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 0.95 \}$

```{r, echo = FALSE, warning = FALSE}
  # This block is a static block that is kept through all 5000 simulations


  # Preparing data
  set.seed(6969)

  # Magic numbers

  dn    <- 100
  dk    <- 10
  va    <- matrix(c(0.3, 0.25, 0.2, 0.15, 0.1, 0.05, 0), nrow = 7, ncol = 1)
  vrho  <- c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 0.95)
  dbeta <- 0
  dsim  <- 100
  rgrid <- seq(0,200,1)

  
  ## INITIALIZATION ##
  # Generate the Z matrix and associates
  mZ   = matrix(rep(0, dn*dk), nrow = dn, ncol = dk)
  
  for (j in 1:ncol(mZ)){
    mZ[,j] <- rnorm(nrow(mZ), mean = 0, sd = 1)
  }
  Pz <- mZ %*% solve(t(mZ) %*% mZ) %*% t(mZ)
  MZ <- diag(nrow=nrow(Pz))-Pz

  # Building a matrix with all Pi
  ve10 <- matrix(c(1, rep(0, dk-1)), nrow=10, ncol = 1)
  mPi  <- t(va %*% t(ve10))

```


```{r, echo = FALSE, warning = FALSE}
# This block simulates the model

fsimul <- function(dn, va, vrho, dBeta, mPi){
  mEV <- array(rep(0, dn*2*length(vrho)), dim = c(dn, 2, length(vrho)))
  
  # Builds the errors sampling from the multivariate normal distribution
  for (i in 1:length(vrho)){
    Sigma   <- matrix(c(1,vrho[i],vrho[i],1),2,2)
    for (j in 1:dn){
      pair    <- mvrnorm(n = 1, rep(0, 2), Sigma)
      mEV[j,,i] <- pair
    }
  }
  
  # Builds X
  lX = list(1,2,3,4,5,6,7,8,9,10)
  
  for(i in 1:length(lX)){
    lX[[i]] <- matrix(rep(0, dn*length(va)), nrow = dn, ncol = length(va))
  }
  
  
  for (j in 1:length(vrho)){
    for (i in 1:length(va)){
      lX[[j]][,i] = mZ %*% mPi[,i] + mEV[,2,j]
    }
  }
  
  # Builds Y
  lY = list(1,2,3,4,5,6,7,8,9,10)
  
  for(i in 1:length(lX)){
    lY[[i]] <- matrix(rep(0, dn*length(va)), nrow = dn, ncol = length(va))
  }
  
  for (j in 1:length(vrho)){
    for (i in 1:length(va)){
      lY[[j]][,i] = lX[[j]][,i] * dbeta + mEV[,1,j]
    }
  }
  
  return(list(lY, lX, mEV))
}
  
```


## Item 1

_For each value of $\alpha$ make a figure of the rejection frequency as a function of $\rho$ when testing $H_0: \beta = 0$ with $95\%$ significance using the $2SLS$ $t$-statistic (so five figures which show the rejection frequency as a function of $\alpha$).

```{r, echo = FALSE, warning = FALSE, eval = FALSE} 
  # First, estimate using 2SLS
  
  ## First stage
  vPihat = list(1,2,3,4,5,6,7,8,9,10)
  
  for(i in 1:length(lX)){
    vPihat[[i]] <- 0*mPi
  }

  
  lXhat = list(1,2,3,4,5,6,7,8,9,10)
  
  for (i in 1:length(lX)){
    lXhat[[i]] <- 0 * lX[[i]]
  }
  

  for (j in 1:length(vrho)){
    for (i in 1:length(va)){
      first_stage     <- lm(lX[[j]][,i] ~ mZ - 1)
      vPihat[[j]][,i] <- first_stage$coefficients
      lXhat[[j]][,i]  <- first_stage$fitted.values
    }
  }
  
  mBetahat <- matrix(rep(0, length(va)*length(vrho)), ncol = length(vrho), nrow = length(va))
  lYhat = list(1,2,3,4,5,6,7,8,9,10)
  
  for (i in 1:length(lY)){
    lYhat[[i]] <- 0 * lY[[i]]
  }
  
  
  # Second stage
  for (j in 1:length(vrho)){
    for (i in 1:length(va)){
      second_stage      <- lm(lY[[j]][,i] ~ lXhat[[j]][,i] - 1)
      mBetahat[i,j]     <- second_stage$coefficients
    }
  }

```

Using a package

```{r, echo = FALSE, warning = FALSE, eval = TRUE}
  simul_results <- vector(mode = "list", length = dsim)
  
  for (k in 1:dsim){
    # Gets the model
    my_model <- list(1,2,3)
    my_model <- fsimul(dn, va, vrho, dBeta, mPi)
    lY       <- my_model[[1]]
    lX       <- my_model[[2]]
    mEV      <- my_model[[3]]

    mBetahat    <- matrix(rep(0, length(va)*length(vrho)), ncol = length(vrho), nrow = length(va))
    mpvaluebeta <- matrix(rep(0, length(va)*length(vrho)), ncol = length(vrho), nrow = length(va))
    mrejectbeta <- matrix(rep(0, length(va)*length(vrho)), ncol = length(vrho), nrow = length(va))
    
    for (j in 1:length(vrho)){
      for (i in 1:length(va)){
        model_2sls         <- ivreg(lY[[j]][,i] ~ lX[[j]][,i] - 1 | mZ[,1] + mZ[,2] + mZ[,3] + mZ[,4] + mZ[,5] + mZ[,6] + mZ[,7] + mZ[,8] + mZ[,9] + mZ[,10])
        mBetahat[i,j]      <- model_2sls$coefficients
        summarybeta        <- summary(model_2sls)
        mpvaluebeta[i,j]   <- summarybeta$coefficients[4]
        mrejectbeta[i,j]   <- ifelse(mpvaluebeta[i,j] < 0.05, 1, 0)
      }
    }
    simul_results[[k]] <- mrejectbeta
  }

  results_final <- 0*simul_results[[1]]

  for (i in 1:dsim){
    results_final <- results_final + simul_results[[i]]
  }

```

Graph for Q1

```{r, echo = FALSE, warning = FALSE, eval = TRUE}

   df1 <- data.frame(Rho = 0, Alpha = 0, Freq = 0)


   for (i in 1:length(va)){
     df  <- data.frame(Rho = vrho, Alpha = va[i], Freq = results_final[i,])
     df1 <- rbind(df1, df)
   }
   df1 <- df1[-1,]
   
   cores <- brewer.pal(7, "Dark2")
   df1$Freq <- round(df1$Freq/dsim*100,0)

   
   p1 <- ggplot(df1[df1$Alpha == 0.30,], aes(x = factor(Rho), y = Freq)) + 
      geom_bar(stat = 'identity', alpha = .4, colour = cores[1], fill = cores[1], width=0.6)+
      scale_x_discrete(TeX('$\\rho$'))+
      scale_y_continuous(TeX('% of times $H_0: \\beta = 0$ was rejected'), breaks = seq(0, 100, by = 10), limits = c(0,100))+
      theme(axis.title = element_text(size=9), axis.text.x = element_text(size = 7), panel.border = element_blank(), axis.line = element_line(colour = "black"))+
      labs(title = TeX('$a = 0.30$'))+
      theme_bw()+
      removeGridX()
      
   
   p2 <- ggplot(df1[df1$Alpha == 0.25,], aes(x = factor(Rho), Freq)) + 
      geom_bar(stat = 'identity', alpha = .4, colour = cores[2], fill = cores[2], width=0.6)+
      scale_x_discrete(TeX('$\\rho$'))+
      scale_y_continuous(TeX('% of times $H_0: \\beta = 0$ was rejected'), breaks = seq(0, 100, by = 10), limits = c(0,100))+
      theme(axis.title = element_text(size=9), axis.text.x = element_text(size = 7), panel.border = element_blank(), axis.line = element_line(colour = "black"))+
      labs(title = TeX('$a = 0.25$'))+
      theme_bw()+
      removeGridX()
   
   p3 <- ggplot(df1[df1$Alpha == 0.20,], aes(x = factor(Rho), Freq)) + 
      geom_bar(stat = 'identity', alpha = .4, colour = cores[3], fill = cores[3], width=0.6)+
      scale_x_discrete(TeX('$\\rho$'))+
      scale_y_continuous(TeX('% of times $H_0: \\beta = 0$ was rejected'), breaks = seq(0, 100, by = 10), limits = c(0,100))+
      theme(axis.title = element_text(size=9), axis.text.x = element_text(size = 7), panel.border = element_blank(), axis.line = element_line(colour = "black"))+
      labs(title = TeX('$a = 0.20$'))+
      theme_bw()+
      removeGridX()
   
   p4 <- ggplot(df1[df1$Alpha == 0.15,], aes(x = factor(Rho), Freq)) + 
      geom_bar(stat = 'identity', alpha = .4, colour = cores[4], fill = cores[4], width=0.6)+
      scale_x_discrete(TeX('$\\rho$'))+
      scale_y_continuous(TeX('% of times $H_0: \\beta = 0$ was rejected'), breaks = seq(0, 100, by = 10), limits = c(0,100))+
      theme(axis.title = element_text(size=9), axis.text.x = element_text(size = 9), panel.border = element_blank(), axis.line = element_line(colour = "black"))+
      labs(title = TeX('$a = 0.15$'))+
      theme_bw()+
      removeGridX()
   
   p5 <- ggplot(df1[df1$Alpha == 0.10,], aes(x = factor(Rho), Freq)) + 
      geom_bar(stat = 'identity', alpha = .4, colour = cores[5], fill = cores[5], width=0.6)+
      scale_x_discrete(TeX('$\\rho$'))+
      scale_y_continuous(TeX('% of times $H_0: \\beta = 0$ was rejected'), breaks = seq(0, 100, by = 10), limits = c(0,100))+
      theme(axis.title = element_text(size=9), axis.text.x = element_text(size = 9), panel.border = element_blank(), axis.line = element_line(colour = "black"))+
      labs(title = TeX('$a = 0.10$'))+
      theme_bw()+
      removeGridX()
   
   p6 <- ggplot(df1[df1$Alpha == 0.05,], aes(x = factor(Rho), Freq)) + 
      geom_bar(stat = 'identity', alpha = .4, colour = cores[6], fill = cores[6], width=0.6)+
      scale_x_discrete(TeX('$\\rho$'))+
      scale_y_continuous(TeX('% of times $H_0: \\beta = 0$ was rejected'), breaks = seq(0, 100, by = 10), limits = c(0,100))+
      theme(axis.title = element_text(size=9), axis.text.x = element_text(size = 9), panel.border = element_blank(), axis.line = element_line(colour = "black"))+
      labs(title = TeX('$a = 0.05$'))+
      theme_bw()+
      removeGridX()
   
   p7 <- ggplot(df1[df1$Alpha == 0.00,], aes(x = factor(Rho), Freq)) + 
      geom_bar(stat = 'identity', alpha = .4, colour = cores[7], fill = cores[7], width=0.6)+
      scale_x_discrete(TeX('$\\rho$'))+
      scale_y_continuous(TeX('% of times $H_0: \\beta = 0$ was rejected'), breaks = seq(0, 100, by = 10), limits = c(0,100))+
      theme(axis.title = element_text(size=9), axis.text.x = element_text(size = 9), panel.border = element_blank(), axis.line = element_line(colour = "black"))+
      labs(title = TeX('$a = 0.00$'))+
      theme_bw()+
      removeGridX()
   
   blank <- grid.rect(gp=gpar(col="white"))
   
    total <- ggplot(df1, aes(x = Rho, y = Freq, color = factor(Alpha))) + 
      geom_line()+
      scale_x_continuous(TeX('$\\rho$'), breaks = seq(0, 1, by = 0.1))+
      scale_y_continuous(TeX('% of times $H_0: \\beta = 0$ was rejected'), breaks = seq(0, 100, by = 10), limits = c(0,100))+
      theme(axis.title = element_text(size=9), axis.text.x = element_text(size = 7), panel.border = element_blank(), axis.line = element_line(colour = "black"))+
      theme_bw()+
      labs(color = TeX("$a$"))+
      removeGridX()
    total <- total+theme(legend.position = c(0.15, 0.55))
    total <- total + theme(legend.background = element_rect(fill="transparent"), legend.text = element_text(size = 6), legend.title = element_text(size=6))

#pdf(file = "C:\\Users\\aisha\\OneDrive\\Documentos\\Mestrado Tinbergen\\Year 2\\Block 02\\Advanced Microeconometrics\\Computer assignment\\Mphil\\AdvMicroeconometrics\\Q1\\Fig01-Q1.pdf", width = 6, height = 3)
  grid.arrange(p1, p2, ncol=2, nrow=1)
#dev.off()  
  
  #pdf(file = "C:\\Users\\aisha\\OneDrive\\Documentos\\Mestrado Tinbergen\\Year 2\\Block 02\\Advanced Microeconometrics\\Computer assignment\\Mphil\\AdvMicroeconometrics\\Q1\\Fig09-Q1.pdf", width = 6, height = 3)
  grid.arrange(p1, p7, ncol=2, nrow=1)
#dev.off()


#pdf(file = "C:\\Users\\aisha\\OneDrive\\Documentos\\Mestrado Tinbergen\\Year 2\\Block 02\\Advanced Microeconometrics\\Computer assignment\\Mphil\\AdvMicroeconometrics\\Q1\\Fig02-Q1.pdf", width = 6, height = 3)
  grid.arrange(p3, p4, ncol=2, nrow=1)
#dev.off()


#pdf(file = "C:\\Users\\aisha\\OneDrive\\Documentos\\Mestrado Tinbergen\\Year 2\\Block 02\\Advanced Microeconometrics\\Computer assignment\\Mphil\\AdvMicroeconometrics\\Q1\\Fig03-Q1.pdf", width = 6, height = 3)
  grid.arrange(p5, p6, ncol=2, nrow=1)
#dev.off()


#pdf(file = "C:\\Users\\aisha\\OneDrive\\Documentos\\Mestrado Tinbergen\\Year 2\\Block 02\\Advanced Microeconometrics\\Computer assignment\\Mphil\\AdvMicroeconometrics\\Q1\\Fig04-Q1.pdf", width = 6, height = 3)
  grid.arrange(p7, blank, ncol=2, nrow=1)
#dev.off()

#pdf(file = "C:\\Users\\aisha\\OneDrive\\Documentos\\Mestrado Tinbergen\\Year 2\\Block 02\\Advanced Microeconometrics\\Computer assignment\\Mphil\\AdvMicroeconometrics\\Q1\\Fig05-Q1.pdf", width = 6, height = 3)
  grid.arrange(p7, total, ncol=2, nrow=1)
#dev.off()


#pdf(file = "C:\\Users\\aisha\\OneDrive\\Documentos\\Mestrado Tinbergen\\Year 2\\Block 02\\Advanced Microeconometrics\\Computer assignment\\Mphil\\AdvMicroeconometrics\\Q1\\Fig06-Q1.pdf", width = 6, height = 3)
  total
#dev.off()
```

## Question 2

Compute and make a figure of the 95% critical value function of the LR statistic as a function of $r(\beta_0)$ for $k=10$. What can you say about the critical value when $r(\beta_0)=0$ or infinite?

```{r, echo = FALSE, warning = FALSE, eval = TRUE}
  
  PsiOne       <- rchisq(dsim, df=1)
  PsiKminusOne <- rchisq(dsim, df=dk-1)
  
  vLR <- matrix(rep(0), dsim*length(rgrid), nrow = dsim, ncol = length(rgrid))
  

  for (g in 1:length(rgrid)){
    for (s in 1:dsim){
      vLR[s,g] <- 0.5*(PsiKminusOne[s] + PsiOne[s] - rgrid[g] +sqrt((PsiKminusOne[s] + PsiOne[s] + rgrid[g])^2-4*rgrid[g]*PsiKminusOne[s]))
    }
  }

  LRcv <- apply(vLR, 2, quantile, probs=c(0.95), na.rm=TRUE)
  df1 <- data.frame(RBeta = rgrid, LRcv)
  
  p1 <- ggplot(df1, aes(x = RBeta, y = LRcv))+
    geom_line(colour = cores[1], size = 0.5)+
    scale_x_continuous(TeX('$r(\\beta_0)$'), limits = c(0, 200))+
    scale_y_continuous(TeX('95% critical value of the LR statistic'), limits = c(0,17), breaks = seq(1,18,1))+
    theme(axis.title = element_text(size=9), axis.text.x = element_text(size = 9), panel.border = element_blank(), axis.line = element_line(colour = "black"))+
    theme_bw()+
    removeGridX()
  
#pdf(file = "C:\\Users\\aisha\\OneDrive\\Documentos\\Mestrado Tinbergen\\Year 2\\Block 02\\Advanced Microeconometrics\\Computer assignment\\Mphil\\AdvMicroeconometrics\\Q1\\Fig07-Q1.pdf", width = 6, height = 3)
  p1
#dev.off()
```

## Question 3

Repeat exercise 1 for the AR, score and LR statistics. What you conclude?

```{r, echo = FALSE, warning = FALSE, eval = TRUE}
 
  simul_resultsAR <- vector(mode = "list", length = dsim)
  crit_AR <- qchisq(0.95,dk)/dk
  mrejectbetaAR <- matrix(rep(0, length(va)*length(vrho)), ncol = length(vrho), nrow = length(va))
  
  
  simul_resultsLM <- vector(mode = "list", length = dsim)
  crit_LM <- qchisq(0.95,1)
  mrejectbetaLM <- matrix(rep(0, length(va)*length(vrho)), ncol = length(vrho), nrow = length(va))
  
  simul_resultsLR <- vector(mode = "list", length = dsim)
  mrejectbetaLR <- matrix(rep(0, length(va)*length(vrho)), ncol = length(vrho), nrow = length(va))
  
  for (k in 1:dsim){
  ### Gets the model
   my_model <- list(1,2,3)
   my_model <- fsimul(dn, va, vrho, dBeta, mPi)
   lY       <- my_model[[1]]
   lX       <- my_model[[2]]
   mEV      <- my_model[[3]]
    
    # For each combination of rho and a, compute the test statistic
    for (j in 1:length(vrho)){
      for (i in 1:length(va)){
        
        dEps               <- lY[[j]][,i] - lX[[j]][,i]*dbeta
        dSigEps            <- (1/(dn-dk))*(t(dEps) %*% MZ %*% (dEps))
        dSigVEps           <- (1/(dn-dk))*(t(dEps) %*% MZ %*% lX[[j]][,i])
        dhatrho            <- dSigVEps/dSigEps
        mTildePi           <- solve(t(mZ) %*% mZ) %*% t(mZ) %*% (lX[[j]][,i] -  dEps * dhatrho)
        ZTildePi           <- mZ %*% mTildePi
        mPZTildePi         <- (ZTildePi) %*% solve(t(ZTildePi) %*% ZTildePi) %*% t(ZTildePi)
        
        # AR
        model_AR_num       <- ((t(dEps) %*% Pz %*% (dEps))/dk)/((t(dEps) %*% MZ %*% (dEps))/(dn-dk))
        mrejectbetaAR[i,j] <- ifelse(model_AR_num > crit_AR, 1, 0)
        
        # LM
        model_LM_num       <- (1/dSigEps)*(t(dEps) %*% mPZTildePi %*% dEps)
        mrejectbetaLM[i,j] <- ifelse(model_LM_num > crit_LM, 1, 0)
        
        # LR
        mSigmaVV   <- (1/(dn-dk))*(t(lX[[j]][,i]) %*% MZ %*% lX[[j]][,i])
        mSigmaVEps <- mSigmaVV - (dSigVEps^2)/dSigEps
        dkAR       <- (1/dSigEps)*(t(dEps) %*% Pz %*% dEps)
        drbeta     <- (1/mSigmaVEps)* (t(mTildePi) %*% t(mZ) %*% mZ %*% mTildePi)
        dLMB0      <- model_LM_num
        
        model_LR_num       <- 0.5*(dkAR - drbeta + sqrt((dkAR + drbeta)^2 - 4*drbeta*(dkAR-dLMB0)))
        cr_pos             <- match(min(abs(rgrid - drbeta)), abs(rgrid - drbeta))
        crit_LR            <- LRcv[cr_pos]
        mrejectbetaLR[i,j] <- ifelse(model_LR_num > crit_LR, 1, 0)
        
      }
    }
    simul_resultsAR[[k]] <- mrejectbetaAR
    simul_resultsLM[[k]] <- mrejectbetaLM
    simul_resultsLR[[k]] <- mrejectbetaLR
  }
  

  ARresults_final <- 0*simul_resultsAR[[1]]

  for (i in 1:dsim){
    ARresults_final <- ARresults_final + simul_resultsAR[[i]]
  }
  
  
  LMresults_final <- 0*simul_resultsLM[[1]]

  for (i in 1:dsim){
    LMresults_final <- LMresults_final + simul_resultsLM[[i]]
  }
  
  LRresults_final <- 0*simul_resultsLR[[1]]

  for (i in 1:dsim){
    LRresults_final <- LRresults_final + simul_resultsLR[[i]]
  }
  
  # Graphs
  
   df1 <- data.frame(Rho = 0, Alpha = 0, Freq = 0)
   df2 <- data.frame(Rho = 0, Alpha = 0, Freq = 0)
   df3 <- data.frame(Rho = 0, Alpha = 0, Freq = 0)


   for (i in 1:length(va)){
     df  <- data.frame(Rho = vrho, Alpha = va[i], Freq = ARresults_final[i,])
     df1 <- rbind(df1, df)
     df  <- data.frame(Rho = vrho, Alpha = va[i], Freq = LRresults_final[i,])
     df2 <- rbind(df2, df)
     df  <- data.frame(Rho = vrho, Alpha = va[i], Freq = LMresults_final[i,])
     df3 <- rbind(df3, df)
   }
   df1 <- df1[-1,]
   df2 <- df2[-1,]
   df3 <- df3[-1,]
   
   cores <- brewer.pal(7, "Dark2")
   df1$Freq <- round(df1$Freq/dsim*100,0)
   df2$Freq <- round(df2$Freq/dsim*100,0)
   df3$Freq <- round(df3$Freq/dsim*100,0)
   
   
    total_AR <- ggplot(df1, aes(x = Rho, y = Freq, color = factor(Alpha))) + 
      geom_line()+
      scale_x_continuous(TeX('$\\rho$'), breaks = seq(0, 1, by = 0.1))+
      scale_y_continuous(TeX('% of times $H_0: \\beta = 0$ was rejected'), breaks = seq(0, 50, by = 5), limits = c(0,50))+
      theme(axis.title = element_text(size=6), axis.text.x = element_text(size = 5), panel.border = element_blank(), axis.line = element_line(colour = "black"))+
      theme_bw()+
      labs(title = TeX('AR'))+
      labs(color = TeX("$a$"))+
      removeGridX()
    total_AR <- total_AR+theme(legend.position = c(0.35, 0.65))
    total_AR <- total_AR + theme(legend.background = element_rect(fill="transparent"), legend.text = element_text(size = 5), legend.title = element_text(size=5), axis.title = element_text(size=6), axis.text.x = element_text(size = 5), axis.text.y = element_text(size = 5), panel.border = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(size = 6), legend.direction="horizontal") 
    
    
    total_LR <- ggplot(df2, aes(x = Rho, y = Freq, color = factor(Alpha))) + 
      geom_line()+
      scale_x_continuous(TeX('$\\rho$'), breaks = seq(0, 1, by = 0.1))+
      scale_y_continuous(TeX('% of times $H_0: \\beta = 0$ was rejected'), breaks = seq(0, 50, by = 5), limits = c(0,50))+
      theme(axis.title = element_text(size=6), axis.text.x = element_text(size = 5), panel.border = element_blank(), axis.line = element_line(colour = "black"))+
      theme_bw()+
      labs(color = TeX("$a$"))+
      labs(title = TeX('LR'))+
      removeGridX()
    total_LR <- total_LR+theme(legend.position = c(0.35, 0.65))
    total_LR <- total_LR + theme(legend.background = element_rect(fill="transparent"), legend.text = element_text(size = 5), legend.title = element_text(size=5), axis.title = element_text(size=6), axis.text.x = element_text(size = 5), axis.text.y = element_text(size = 5), panel.border = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(size = 6), legend.direction="horizontal") 

    
    total_LM <- ggplot(df3, aes(x = Rho, y = Freq, color = factor(Alpha))) + 
      geom_line()+
      scale_x_continuous(TeX('$\\rho$'), breaks = seq(0, 1, by = 0.1))+
      scale_y_continuous(TeX('% of times $H_0: \\beta = 0$ was rejected'), breaks = seq(0, 50, by = 5), limits = c(0,50))+
      theme(axis.title = element_text(size=6), axis.text.x = element_text(size = 5), axis.text.y = element_text(size = 5), panel.border = element_blank(), axis.line = element_line(colour = "black"))+
      theme_bw()+
      labs(color = TeX("$a$"))+
      labs(title = TeX('LM'))+
      removeGridX()
    total_LM <- total_LM+theme(legend.position = c(0.35, 0.65))
    total_LM <- total_LM + theme(legend.background = element_rect(fill="transparent"), legend.text = element_text(size = 5), legend.title = element_text(size=5), axis.title = element_text(size=6), axis.text.x = element_text(size = 5), axis.text.y = element_text(size = 5), panel.border = element_blank(), axis.line = element_line(colour = "black"), plot.title = element_text(size = 6), legend.direction="horizontal") 

#pdf(file = "C:\\Users\\aisha\\OneDrive\\Documentos\\Mestrado Tinbergen\\Year 2\\Block 02\\Advanced Microeconometrics\\Computer assignment\\Mphil\\AdvMicroeconometrics\\Q1\\Fig08-Q1.pdf", width = 9, height = 3)
  grid.arrange(total_AR, total_LR, total_LM, ncol=3, nrow=1) 
#dev.off() 

``` 